---
title: "CEJS Activity"
format: html
---

<!-- solution-begin -->
These are instructor notes; they will be removed from the student-facing assignment file.

This is the **Bayesian** form of the activity.
<!-- solution-end -->

```{r setup}
#| include: false
#| echo: false
library(tidyverse)
library(rstanarm)
library(bayesplot)
library(broom.mixed)

filename <- "../data/1.0-communities.csv"
```

```{r wrangling}
#| include: false
#| echo: false
df_raw <- read_csv(filename)
df_data <- 
  df_raw %>%
  janitor::clean_names()
```

# Overview

(Preview the overview figure)

```{r fig-overview}
#| warning: false
#| echo: false
#| fig-align: center
#| fig-height: 8
df_data %>% 
  filter(state_territory %in% c("Florida", "Massachusetts", "Colorado")) %>% 
  
  ggplot(aes(
    percent_black_or_african_american_alone,
    energy_burden_percentile
  )) +
  geom_point(size = 0.1) +
  geom_smooth(method = "lm") +
  facet_wrap(~state_territory, ncol = 2) +
  scale_x_continuous(labels = scales::percent) +
  theme_minimal() +
  labs(
    x = "Percent Black Census Tract",
    y = "Energy Burden Percentile",
  )
```

- Florida and Massachusetts have a clear positive trend; this is evidence of inequitable outcomes
- Colorado has a negative trend; this is a surprising result

# Analyze a Bayesian Model

(Introduce the Bayesian model)

\[ B = m P + b, \]

where $B$ is the energy burden percentile, $P$ is the percent black, $m$ is the slope parameter, and $b$ is the intercept parameter.

```{r fit-MA}
#| include: false
#| echo: false
#| cache: true
model_MA <- stan_glm(
  data = df_data %>% 
    filter(
      str_detect(state_territory, "Massachusetts"),
      !is.na(percent_black_or_african_american_alone)
    ),
  # Likelihood model
  formula = energy_burden_percentile ~ percent_black_or_african_american_alone,
  family = gaussian,           # link function
  # Prior model, use the Stan defaults
  # Numerical parameters
  chains = 4, 
  iter = 5000*2, # to account for 50% burn-in
  seed = 84735,  # the BAYES seed!
  refresh = FALSE
) 
```

## Study the posterior distribution

(Introduce the posterior)

```{r posteriors-MA}
#| warning: false
#| echo: false
#| fig-align: center
#| fig-height: 8
df_model_MA <- 
  model_MA %>% 
  as_tibble() %>% 
  select(Intercept = `(Intercept)`, Slope = percent_black_or_african_american_alone)

df_model_MA %>% 
  pivot_longer(
    c(Intercept, Slope),
    names_to = "parameter",
    values_to = "x"
  ) %>% 
  
  ggplot(aes(x)) +
  geom_density() +
  facet_wrap(~parameter) +
  theme_minimal() +
  labs(
    x = "Parameter Value",
    y = "Density",
  )
```

::: {.callout-note icon=false title="Synthesizing Descriptive Statistics"}
Which model parameter describes the **trend** of the energy burden? What does the posterior say about that parameter value?
:::

The posterior distribution helps us determine **how confident** we should be in conclusions drawn from the model.

## Assessing confidence

TODO teach students how to do this through an example

## Study the posterior predictions

(Describe posterior predictions)

```{r}
#| warning: false
#| echo: false
#| fig-align: center
#| fig-height: 8
df_data %>% 
  filter(str_detect(state_territory, "Massachusetts")) %>% 
  
  ggplot(aes(
    x = percent_black_or_african_american_alone,
    y = energy_burden_percentile
  )) +
  geom_line(
    data = . %>% 
      filter(!is.na(percent_black_or_african_american_alone)) %>% 
      select(percent_black_or_african_american_alone) %>% 
      tidybayes::add_fitted_draws(model_MA, n = 1000),
    aes(y = .value, group = .draw), 
    alpha = 0.01,
    linewidth = 2.0,
    color = "blue"
  ) +
  geom_point(size = 0.1) +
  scale_x_continuous(labels = scales::percent) +
  theme_minimal() +
  labs(
    x = "Percent Black Census Tract",
    y = "Energy Burden Percentile",
  )
```

- How well do the posterior predictions agree with the data?

# Colorado

## Set a prior

(Which of the following best matches the MA posterior?)

```{r}
#| warning: false
#| echo: false
#| fig-align: center
#| fig-height: 8
mean_opt1 <- 10
sd_opt1 <- 4.7
mean_opt2 <- 52
sd_opt2 <- 4.7
mean_opt3 <- 52
sd_opt3 <- 8.2
  
tibble(Slope = seq(-10, 80, length.out = 200)) %>% 
  mutate(
    d_1 = dnorm(Slope, mean = mean_opt1, sd = sd_opt1),
    d_2 = dnorm(Slope, mean = mean_opt2, sd = sd_opt2),
    d_3 = dnorm(Slope, mean = mean_opt3, sd = sd_opt3),
  ) %>% 
  pivot_longer(
    cols = matches("d_\\d"),
    names_sep = "_",
    names_to = c(".value", "Option")
  ) %>% 

  ggplot(aes(Slope)) +
  geom_density(
    data = df_model_MA,
    mapping = aes(y = ..density..),
  ) +
  geom_line(aes(y = d, color = Option)) +
  facet_grid(~Option, labeller = label_both) +
  theme_minimal() +
  guides(color = "none") +
  labs(
    y = "Density"
  )
```

## Study the posterior predictions

```{r}
#| include: false
#| echo: false
#|cache: true
model_CO1 <- stan_glm(
  data = df_data %>% 
    filter(
      str_detect(state_territory, "Colorado"),
      !is.na(percent_black_or_african_american_alone)
    ),
  # Likelihood model
  formula = energy_burden_percentile ~ percent_black_or_african_american_alone,
  family = gaussian,           # link function
  # Use parameter set 1
  prior = normal(mean_opt1, sd_opt1),
  # Use an "overconfident" prior for the intercept, based on the energy burden posterior for MA
  prior_intercept = normal(51, 1),
  # Numerical parameters
  chains = 4, 
  iter = 5000*2, # to account for 50% burn-in
  seed = 84735,  # the BAYES seed!
  refresh = FALSE
) 

model_CO2 <- stan_glm(
  data = df_data %>% 
    filter(
      str_detect(state_territory, "Colorado"),
      !is.na(percent_black_or_african_american_alone)
    ),
  # Likelihood model
  formula = energy_burden_percentile ~ percent_black_or_african_american_alone,
  family = gaussian,           # link function
  # Use parameter set 2
  prior = normal(mean_opt2, sd_opt2),
  # Use an "overconfident" prior for the intercept, based on the energy burden posterior for MA
  prior_intercept = normal(51, 1),
  # Numerical parameters
  chains = 4, 
  iter = 5000*2, # to account for 50% burn-in
  seed = 84735,  # the BAYES seed!
  refresh = FALSE
) 

model_CO3 <- stan_glm(
  data = df_data %>% 
    filter(
      str_detect(state_territory, "Colorado"),
      !is.na(percent_black_or_african_american_alone)
    ),
  # Likelihood model
  formula = energy_burden_percentile ~ percent_black_or_african_american_alone,
  family = gaussian,           # link function
  # Use parameter set 3
  prior = normal(mean_opt3, sd_opt3),
  # Use an "overconfident" prior for the intercept, based on the energy burden posterior for MA
  prior_intercept = normal(51, 1),
  # Numerical parameters
  chains = 4, 
  iter = 5000*2, # to account for 50% burn-in
  seed = 84735,  # the BAYES seed!
  refresh = FALSE
) 
```


```{r}
#| warning: false
#| echo: false
#| fig-align: center
#| fig-height: 8
df_data %>% 
  filter(str_detect(state_territory, "Colorado")) %>% 
  
  ggplot(aes(
    x = percent_black_or_african_american_alone,
    y = energy_burden_percentile
  )) +
  geom_line(
    data = . %>% 
      filter(!is.na(percent_black_or_african_american_alone)) %>% 
      select(percent_black_or_african_american_alone) %>% 
      tidybayes::add_fitted_draws(model_CO1, n = 1000) %>% 
      mutate(Option = "1"),
    aes(y = .value, group = .draw, color = Option), 
    alpha = 0.01,
    linewidth = 2.0,
  ) +
  geom_line(
    data = . %>% 
      filter(!is.na(percent_black_or_african_american_alone)) %>% 
      select(percent_black_or_african_american_alone) %>% 
      tidybayes::add_fitted_draws(model_CO2, n = 1000) %>% 
      mutate(Option = "2"),
    aes(y = .value, group = .draw, color = Option), 
    alpha = 0.01,
    linewidth = 2.0,
  ) +
  geom_line(
    data = . %>% 
      filter(!is.na(percent_black_or_african_american_alone)) %>% 
      select(percent_black_or_african_american_alone) %>% 
      tidybayes::add_fitted_draws(model_CO3, n = 1000) %>% 
      mutate(Option = "3"),
    aes(y = .value, group = .draw, color = Option), 
    alpha = 0.01,
    linewidth = 2.0,
  ) +
  
  geom_point(size = 0.1) +
  facet_grid(~Option, labeller = label_both) +
  scale_x_continuous(labels = scales::percent) +
  theme_minimal() +
  guides(color = "none") +
  labs(
    x = "Percent Black Census Tract",
    y = "Energy Burden Percentile",
    title = "Colorado"
  )
```

::: {.callout-note icon=false title="Synthesizing Descriptive Statistics"}
What does the model suggest about the trend of energy burden with percent Black in Colorado? How confident are you in your conclusion?
:::

<!-- solution-begin -->
The posterior predictions do not match the data well, due to the overconfident prior. This is a contrast with the Frequentist approach.
<!-- solution-end -->

# Florida 

## Deal with Limited Data

(Spin a yarn about how FL might redact their data)

```{r}
#| warning: false
#| echo: false
#| fig-align: center
#| fig-height: 8
set.seed(101)

df_fl <- 
  df_data %>% 
  filter(state_territory == "Florida") %>% 
  slice_sample(n = 25)

df_fl %>% 
  ggplot(aes(
    percent_black_or_african_american_alone,
    energy_burden_percentile
  )) +
  geom_point(size = 1.0) +
  geom_smooth(method = "lm") +
  scale_x_continuous(labels = scales::percent) +
  theme_minimal() +
  labs(
    x = "Percent Black Census Tract",
    y = "Energy Burden Percentile",
  )
```

- Given the limited data, we can't confidently conclude that there is a positive slope

## Study the posterior predictions

```{r}
#| include: false
#| echo: false
#|cache: true
model_FL1 <- stan_glm(
  data = df_fl,
  # Likelihood model
  formula = energy_burden_percentile ~ percent_black_or_african_american_alone,
  family = gaussian,           # link function
  # Use parameter set 1
  prior = normal(mean_opt1, sd_opt1),
  # Use an "overconfident" prior for the intercept, based on the energy burden posterior for MA
  prior_intercept = normal(51, 1),
  # Numerical parameters
  chains = 4, 
  iter = 5000*2, # to account for 50% burn-in
  seed = 84735,  # the BAYES seed!
  refresh = FALSE
) 

model_FL2 <- stan_glm(
  data = df_fl,
  # Likelihood model
  formula = energy_burden_percentile ~ percent_black_or_african_american_alone,
  family = gaussian,           # link function
  # Use parameter set 2
  prior = normal(mean_opt2, sd_opt2),
  # Use an "overconfident" prior for the intercept, based on the energy burden posterior for MA
  prior_intercept = normal(51, 1),
  # Numerical parameters
  chains = 4, 
  iter = 5000*2, # to account for 50% burn-in
  seed = 84735,  # the BAYES seed!
  refresh = FALSE
) 

model_FL3 <- stan_glm(
  data = df_fl,
  # Likelihood model
  formula = energy_burden_percentile ~ percent_black_or_african_american_alone,
  family = gaussian,           # link function
  # Use parameter set 3
  prior = normal(mean_opt3, sd_opt3),
  # Use an "overconfident" prior for the intercept, based on the energy burden posterior for MA
  prior_intercept = normal(51, 1),
  # Numerical parameters
  chains = 4, 
  iter = 5000*2, # to account for 50% burn-in
  seed = 84735,  # the BAYES seed!
  refresh = FALSE
) 
```


```{r}
#| warning: false
#| echo: false
#| fig-align: center
#| fig-height: 8
df_fl %>% 
  ggplot(aes(
    x = percent_black_or_african_american_alone,
    y = energy_burden_percentile
  )) +
  geom_line(
    data = . %>% 
      filter(!is.na(percent_black_or_african_american_alone)) %>% 
      select(percent_black_or_african_american_alone) %>% 
      tidybayes::add_fitted_draws(model_FL1, n = 1000) %>% 
      mutate(Option = "1"),
    aes(y = .value, group = .draw, color = Option), 
    alpha = 0.01,
    linewidth = 2.0,
  ) +
  geom_line(
    data = . %>% 
      filter(!is.na(percent_black_or_african_american_alone)) %>% 
      select(percent_black_or_african_american_alone) %>% 
      tidybayes::add_fitted_draws(model_FL2, n = 1000) %>% 
      mutate(Option = "2"),
    aes(y = .value, group = .draw, color = Option), 
    alpha = 0.01,
    linewidth = 2.0,
  ) +
  geom_line(
    data = . %>% 
      filter(!is.na(percent_black_or_african_american_alone)) %>% 
      select(percent_black_or_african_american_alone) %>% 
      tidybayes::add_fitted_draws(model_FL3, n = 1000) %>% 
      mutate(Option = "3"),
    aes(y = .value, group = .draw, color = Option), 
    alpha = 0.01,
    linewidth = 2.0,
  ) +
  
  geom_point(size = 0.1) +
  facet_grid(~Option, labeller = label_both) +
  scale_x_continuous(labels = scales::percent) +
  theme_minimal() +
  guides(color = "none") +
  labs(
    x = "Percent Black Census Tract",
    y = "Energy Burden Percentile",
    title = "Florida, Limited Data"
  )
```

::: {.callout-note icon=false title="Synthesizing Descriptive Statistics"}
What does the model suggest about the trend of energy burden with percent Black in Florida? How confident are you in your conclusion?
:::

<!-- solution-begin -->
- With the prior, we can confidently conclude there is a positive slope
<!-- solution-end -->
