---
title: "Progress Report 1: Bayes BATS Tier-2"
subtitle: "Activity: Comparing Frequentist and Bayesian Approaches"
format: html
author: 
  - name: Stefani Langehennig
  - name: Zach del Rosario
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

library(plan)
library(dplyr)
```

## Overview

Our aim for Progress Report 1 was to complete a number of "starter" artifacts ahead of creating what we believe is the heart of our activity - an application in `R` showing the comparison between Bayesian and frequentist hypothesis testing. Our rationale for doing so was simple: having a dataset in hand, alongside some context for that dataset, would set the stage for the application to come. As well, we created introductory and concluding discussion artifacts, as these helped frame our own thinking ahead of implementing our applied comparison activity using Bayesian and frequentist approaches.

## Updated Gantt Chart

Below is the updated Gantt chart we proposed as part of our overall work plan. To date, we completed progress on the artifacts we planned to for the submission of Progress Report 1. Links to those artifacts are provided below.

```{r, echo=FALSE}
g <- new("gantt")
g <- ganttAddTask(g, "Work Plan")
g <- ganttAddTask(g, "ID datasets","2023-08-09","2023-08-25", done=100)
g <- ganttAddTask(g, "Draft context 1-pager","2023-08-25","2023-09-01",done=90)
g <- ganttAddTask(g, "Progress Report 1")
g <- ganttAddTask(g, "Finalize context 1-pager", "2023-09-01", "2023-09-30", done=50)
g <- ganttAddTask(g, "Draft intro. discussion", "2023-09-01", "2023-09-22", done=75)
g <- ganttAddTask(g, "Finalize intro. discussion", "2023-09-22", "2023-09-30", done=25)
g <- ganttAddTask(g, "Draft conclusion discussion", "2023-09-01", "2023-09-22", done=25)
g <- ganttAddTask(g, "Finalize conclusion discussion", "2023-09-22", "2023-09-30", done=0)
g <- ganttAddTask(g, "Draft Bayes-freq. framework 1-pager", "2023-09-01", "2023-09-30", done=75)
g <- ganttAddTask(g, "Progress Report 2")
g <- ganttAddTask(g, "Finalize Bayes-freq. framework 1-pager", "2023-10-01", "2023-10-20", done=0)
g <- ganttAddTask(g, "Draft applied R activity", "2023-10-01", "2023-10-31", done=25)
g <- ganttAddTask(g, "Final Materials Submission")
g <- ganttAddTask(g, "Finalize applied R activity", "2023-11-01", "2023-11-24", done=0)
g <- ganttAddTask(g, "Finalize all student artifacts", "2023-11-01", "2023-11-24", done=0)
g <- ganttAddTask(g, "Develop video presentation", "2023-11-24", "2023-11-30", done=0)
g <- ganttAddTask(g, "Conference prep", "2024-01-05", "2024-02-24", done=0)


font <- ifelse(is.na(g[["start"]]), 2, 1)

plot(g, ylabels=list(font=font),
     event.label='Submit Final Materials',event.time='2023-12-01',
     col.event=c("firebrick2"),
     col.done=c("steelblue3"),
     col.notdone=c("palegreen3"),
     cex.main=1,
     main="Tier 2: Freq-Bayes Gantt Chart",
     )

legend("topright", pch=22, pt.cex=2, cex=0.6, pt.bg=c("steelblue3", "palegreen3"),
     border="black",
     legend=c("Completed", "Incomplete"), bg="white", xpd=TRUE)

```

## Completed Artifacts

* [Dataset: The Climate and Economic Justice Screening Tool (CEJST)](https://screeningtool.geoplatform.gov/en/)
* [Context Reference Guide](https://github.com/bayes-bats/tier2-freq-bayes/blob/main/development/context.qmd)
* [Critical Differences Reference Guide](https://github.com/bayes-bats/tier2-freq-bayes/blob/main/development/one-pager.qmd)
* [Introduction Discussion](https://github.com/bayes-bats/tier2-freq-bayes/blob/main/development/introduction.qmd)
* [Conclusion Discussion](https://github.com/bayes-bats/tier2-freq-bayes/blob/main/development/conclusion.qmd)


## Next Steps

As we move into the second phase of creating our activity and its associated artifacts, the primary area of focus will be on drafting the applied `R` activity. We are currently in the process of developing a [series of prototypes](https://github.com/bayes-bats/tier2-freq-bayes/blob/main/development/communities.qmd) using the CEJST data. Our hope is to build a [main activity](https://github.com/bayes-bats/tier2-freq-bayes/blob/main/development/activity-main.qmd) based on the prototyping that we will then finalize in the final phase of our activity development.

Once we finalize which prototype we want to use for our main activity, we will refine it so that it is ready for student consumption. In addition to drafting the applied `R` activity, we will also finalize the [Critical Differences Reference Guide](https://github.com/bayes-bats/tier2-freq-bayes/blob/main/development/one-pager.qmd) so that it aligns with the application portion of our activity. We anticipate having both of these steps (drafting the applied activity and finalizing the differences reference guide) complete by the time Progress Report 2 is due on 01 November.

## Meeting Times

The cadence for meetings has not changed between our work plan submission and Progress Report 1, thus we will keep meeting based on the following:

-   Weekly and/or bi-weekly meetings for the activity-creating team
-   Monthly meetings with mentor (Mine)
-   Office hours
    -   03 October 2023 with Monika (10-11a EST)
    -   16 October 2023 with Amy (2-3p EST)
    -   26 October 2023 with Mine (1-2p EST)
    -   November (TBD)

