---
title: "Progress Report 2: Bayes BATS Tier 2"
subtitle: "Activity: Comparing Frequentist and Bayesian Approaches"
format: html
author: 
  - name: Stefani Langehennig
  - name: Zach del Rosario
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

library(plan)
library(dplyr)
```

## Overview
The second phase of creating our activity focused primarily on drafting the applied `R` activity. We began building the framework for the applied activity ahead of Progress Report 1 and have made considerable progress on it since then. Links to the applied activity and other artifacts are below in @sec-artifacts, **Artifacts**.

Highlights from this phase:
- Completion of both the Context and Introductory documents, which now include clear learning objectives, instructor solutions/guidelines, and more information on statistical inference
- 


## Updated Gantt Chart

```{r, echo=FALSE}
g <- new("gantt")
g <- ganttAddTask(g, "Work Plan")
g <- ganttAddTask(g, "ID datasets","2023-08-09","2023-08-25", done=100)
g <- ganttAddTask(g, "Draft context 1-pager","2023-08-25","2023-09-01",done=100)
g <- ganttAddTask(g, "Progress Report 1")
g <- ganttAddTask(g, "Finalize context 1-pager", "2023-09-01", "2023-09-30", done=100)
g <- ganttAddTask(g, "Draft intro. discussion", "2023-09-01", "2023-09-22", done=100)
g <- ganttAddTask(g, "Finalize intro. discussion", "2023-09-22", "2023-09-30", done=100)
g <- ganttAddTask(g, "Draft conclusion discussion", "2023-09-01", "2023-09-22", done=25)
g <- ganttAddTask(g, "Draft Bayes-freq. framework 1-pager", "2023-09-01", "2023-09-30", done=100)
g <- ganttAddTask(g, "Submit Tier 3 Proposal", "2023-09-23", "2023-09-30", done=100)
g <- ganttAddTask(g, "Progress Report 2")
g <- ganttAddTask(g, "Finalize Bayes-freq. framework 1-pager", "2023-10-01", "2023-10-20", done=50)
g <- ganttAddTask(g, "Draft applied R activity", "2023-10-01", "2023-10-31", done=85)
g <- ganttAddTask(g, "Finalize conclusion discussion", "2023-09-22", "2023-09-30", done=0)
g <- ganttAddTask(g, "Final Materials Submission")
g <- ganttAddTask(g, "Finalize applied R activity", "2023-11-01", "2023-11-24", done=0)
g <- ganttAddTask(g, "Finalize all student artifacts", "2023-11-01", "2023-11-24", done=0)
g <- ganttAddTask(g, "Develop video presentation", "2023-11-24", "2023-11-30", done=0)
g <- ganttAddTask(g, "Conference prep", "2024-01-05", "2024-02-24", done=0)


font <- ifelse(is.na(g[["start"]]), 2, 1)

plot(g, ylabels=list(font=font),
     event.label='Submit Final Materials',event.time='2023-12-01',
     col.event=c("firebrick2"),
     col.done=c("steelblue3"),
     col.notdone=c("palegreen3"),
     cex.main=1,
     main="Tier 2: Freq-Bayes Gantt Chart",
     )

legend("topright", pch=22, pt.cex=2, cex=0.6, pt.bg=c("steelblue3", "palegreen3"),
     border="black",
     legend=c("Completed", "Incomplete"), bg="white", xpd=TRUE)

```

## Artifacts {#sec-artifacts}

###### Completed
* [Critical Differences Reference Guide](https://github.com/bayes-bats/tier2-freq-bayes/blob/main/development/03-one-pager-main.qmd)
* [Conclusion Discussion](https://github.com/bayes-bats/tier2-freq-bayes/blob/main/development/04-conclusion-main.qmd)

###### In-Progress
* [Applied Activity using the Bayesian approach](https://github.com/bayes-bats/tier2-freq-bayes/blob/main/development/02b-activity-bayes-main.qmd)
* [Applied Activity using the Frequentist approach](https://github.com/bayes-bats/tier2-freq-bayes/blob/main/development/02a-activity-freq-main.qmd)

## Next Steps



In our monthly meeting (2023-10-19), Mine gave us the following feedback on the activity, which we will incorporate in the next phase of work:


## Meeting Times

