---
title: "Surprise - They’re Different!"
subtitle: "Comparing Frequentist and Bayesian Approaches in Public Policy"
author: "Stefani Langehennig, Zach del Rosario, Mine Dogucu"
institute: "2024 Midwest Political Science Association Conference"
date: "06 April 2024"
output:
  xaringan::moon_reader:
    css: [xaringan-themer.css, "ninjutsu"]
    lib_dir: libs
    nature:
      highlightStyle: solarized-light
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent(base_color = "#2F4F4F",
  header_font_google = google_font("Yanone Kaffeesatz","200","300","400","500","600","700"),
  text_font_google   = google_font("Roboto", "300", "300i","400","500"),
  code_font_google   = google_font("Fira Mono"),
  text_font_size = "125%", 
  colors = c(
  red = "#f34213",
  purple = "#3e2f5b",
  orange = "#ff8811",
  green = "#136f63",
  white = "#FFFFFF",
  gold = "#FFD700",
  crimson = "#8B0000"
),
extra_css = list(
    ".remark-slide table" = list("display" = "table",
                   "width" = "80%",
                   "text-align" = "left"),
    ".remark-slide-number" = list("display" = "%current%"),
    ".strong" = list("font-weight" = "400"),
    ".big" = list("font-size" = "350%",
                  "font-family" = "Yanone Kaffeesatz",
                  "font-weight"="400"),
    ".small" = list("font-size" = "80%"),
    ".large" = list("font-size" = "150%"))
)

```

# Agenda
- Inferential Statistics & Public Policy
- Applied Activity: Comparing Frequentist & Bayesian Approaches
- Evaluation
- Conclusion

---

class: left, bottom, inverse
.big[Inferential Statistics & Public Policy]

---

# Inferential Statistics & Public Policy

- here

---

class: left, bottom, inverse
.big[Applied Activity: Comparing Frequentist & Bayesian Approaches]

---

# Applied Activity: Comparing Frequentist & Bayesian Approaches

- Problem context
- Activity Application
- Activity Closing

---


class: left, bottom, inverse
.big[Evaluation]

---

# Evaluation

- survey design
- results

---


class: left, bottom, inverse
.big[Conclusions]

---

# Conclusions

- implications, directions for future research, etc.

---

class: left, bottom, inverse
.big[Thank you! Questions?]